### 2015-07-04 narrative/placement.tex
reorged the placement subsubsection

Our past work on data
staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
in HPC systems, coupled with the experience in developing high performance
I/O transports for ADIOS \cite{liu2014hello,lofstead2008flexible} has
provided us with insights on how to manage the complex I/O requirements for
exascale applications, particularly in the face of the aforementioned
multi-tier storage hierarchy. A key requirements is application-driven
%
runtime mechanisms for dynamically managing data placement across the layers
of the distributed storage hierarchy,
%
coordinating data movement 
%
and data sharing between the components of the application workflow so as to
maximize its utility to the application and reduce access costs. Our
approach to data placement will leverage the increased set of knowledge
about the data available to the middleware and storage system, potentially
allowing for greater optimizations as compared to other storage solutions.

@@

\paragraph{Proposed research approach:} In this project we will utilize
deep memory hierarchies and multilevel storage structures to support
staging-based in-transit data management and processing for data-intensive
scientific workflows at extreme scale. Specifically, we will address the
coordination and data access/sharing requirements of application workflows,
as well as the online data management requirements such as for example,
in-transit online data analysis, check-pointing, data reduction. We will
also deal with the complexities associated with data placement of data and
movement in the multiple layers and components of the storage hierarchy. The
key components of our research include:

{\em Application hints:} Application hints can provide additional
information to the runtime based on the users' knowledge of the application
workflow as well as on past experiences, for example, possible data access
patterns, nature of regions of interest, and domain specific information.
This application level information will be added to the specification of the
workflow and can provide advisory inputs to efficiently help the runtime
make smarter decisions about the placement and/or movement of data objects
and the scheduling and execution of tasks, as well as to effectively manage
trade-offs between for example, power consumption and time-to-solution.

{\em Data utility:} We will formulate the concept of utility (from an
application perspective) that is assigned data objects to quantify the
benefits of placement and/or movement actions to application. These
utilities may be based on the importance of the data objects, the frequency
with which they are accessed, etc., and are used during runtime data
management. For example, data objects will higher utility value may be
placed closer (e.g., in faster memory) to the accessing application.
Similarly, when it is time to evict data objects from a higher level storage
layer, the decision about which object to evict may take the utility into
account.

{\em Autonomic in-transit data management:} Efficiently placing data objects
vertically across multiple memory hierarchy layers and storage levels as
well as horizontally across different storage nodes requires balancing
competing objectives. In this project we will explore autonomic data
management strategies that can evaluate utility/cost tradeoffs and
appropriately place/move data objects at runtime. The autonomic approach
will leverage both user hints and information gathered at runtime (e.g.,
runtime data access history, network topology, etc.) to place data object.

{\em Runtime tracking and estimation:} Autonomic data placement and/or
movement will leverage the information gathered at runtime together with
predictive analysis requirements. This information may include runtime
spatial-temporal data access patterns, physical network topology
information, and etc., which will be tracked and gathered at runtime, and
then be used to estimate and predict the behavior of coupled applications in
the rest of simulation-time. A key research aspect here will be to
anticipate accesses, for example using learning techniques or using
application knowledge, and use this information to prefetch data objects.
For example, we will identify and catalogue key phases and access patterns
in the target workflows and then use it to characterize patterns at runtime
and trigger appropriate place/movement actions.

>>>>

Our past work on data
staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
in HPC systems, coupled with the experience in developing high performance
I/O transports for ADIOS \cite{liu2014hello,lofstead2008flexible} has
provided us with insights on how to manage the complex I/O requirements for
exascale applications, particularly in the face of the aforementioned
multi-tier storage hierarchy. A key requirements is application-driven
%
runtime mechanisms for dynamically managing data placement across the layers
of the distributed storage hierarchy,
%
coordinating data movement 
%
and data sharing between the components of the application workflow so as to
maximize its utility to the application and reduce access costs. Our
approach to data placement will leverage the increased set of knowledge
about the data available to the middleware and storage system, potentially
allowing for greater optimizations as compared to other storage solutions. 
The key components of our research include

\paragraph{Application Hints:}
As described in Section~\ref{sec:refactor} one of the main components of our
proposed storage system is the ability to refactor and reduce data as it is
generated, and to reorgnized and regenerate the data as it is accessed. We
carry this principle into the placement and movement of data by allowing
applications to define hints and policies that guide \textbf{what} data is
placed \textbf{where}. We will explore the use of application hints in two
distinct areas. Firstly, we will study the challenges and tradeoffs of either
augmenting the I/O interface with hints or allowing the additional of an
external specification that defines the use case. Our experience with
developing modern I/O interfaces has shown that both techniques have value,
and we will investigate the set of hints that are embedded in the
application code vs. those that are described within an non-compiled
specification. Secondly, we will study how hints can guide the initial data
placement as data is handed off from application to storage ( during a
write) and from storage to application (during the read). In both cases we
will study what minimal set of annotations and hints can allow the storage
system to minimize data movement and optimize the resources consumed by I/O.

\paragraph{Data Utility:} We will formulate the concept of utility (from an
application perspective) that is assigned data objects to quantify the
benefits of placement and/or movement actions to application. These
utilities may be based on the importance of the data objects, the frequency
with which they are accessed, etc., and are used during runtime data
management. For example, data objects will higher utility value may be
placed closer (e.g., in faster memory) to the accessing application.
Similarly, when it is time to evict data objects from a higher level storage
layer, the decision about which object to evict may take the utility into
account.

\paragraph{Autonomic in-transit data management:} 
Efficiently placing data objects
vertically across multiple memory hierarchy layers and storage levels as
well as horizontally across different storage nodes requires balancing
competing objectives. In this project we will explore autonomic data
management strategies that can evaluate utility/cost tradeoffs and
appropriately place/move data objects at runtime. The autonomic approach
will leverage both user hints and information gathered at runtime (e.g.,
runtime data access history, network topology, etc.) to place data object.

\paragraph{ Runtime tracking and estimation:}
Autonomic data placement and/or
movement will leverage the information gathered at runtime together with
predictive analysis requirements. This information may include runtime
spatial-temporal data access patterns, physical network topology
information, and etc., which will be tracked and gathered at runtime, and
then be used to estimate and predict the behavior of coupled applications in
the rest of simulation-time. A key research aspect here will be to
anticipate accesses, for example using learning techniques or using
application knowledge, and use this information to prefetch data objects.
For example, we will identify and catalogue key phases and access patterns
in the target workflows and then use it to characterize patterns at runtime
and trigger appropriate place/movement actions.

@@
